---
title: "Building dataframes from the economic calendar of Forex Factory"
output: html_document
date: "2023-02-02"
---

# Loading R packages
```{r}
library(dplyr)
library(rvest)
library(lubridate)
library(stringr)
library(zoo)
library(magrittr)
library(tidyr)
library(RSelenium)
```

# Building a partial dataset excluding Forex Factory's details variables

## Scrapping data from Forex Factory 
```{r}
date_range = seq(as.Date("2007-01-01"),as.Date("2022-02-06"), by = "days")

df_FF = c()

prev_month = as.yearmon(date_range[1])

for (i in 1:length(date_range)) {
  day_ref = day(date_range[i])
  month_ref = str_to_lower(month(date_range[i], label = TRUE))
  year_ref = year(date_range[i])
  
  date_ref = paste0(month_ref, day_ref, ".", year_ref)
  
  page_link = paste0("https://www.forexfactory.com/calendar?day=", date_ref)
  
  elements <- read_html(page_link) %>% 
    html_elements(".calendar__table") %>% 
    html_elements(".calendar__row--grey")
  
  currency <- elements %>%
    html_elements(".currency") %>%
    html_text2()
  
  event <- elements %>%
    html_elements(".event") %>%
    html_text2()
  
  forecast <- elements %>%
    html_elements(".forecast") %>%
    html_text2()
  
  observation <- elements %>%
    html_elements(".actual") %>%
    html_text2()
  
  previous_obs <- elements %>%
    html_elements(".previous") %>%
    html_text2()
  
  impact = elements %>%
    html_elements(".calendar__impact-icon--screen") %>%
    html_elements("[title]") %>%
    html_attr("title")
  
  event_id = elements %>%
    html_attr("data-eventid")
  
  df_day = tibble(currency,
                  event,
                  forecast,
                  observation,
                  previous_obs,
                  impact,
                  event_id) %>%
    mutate(date = date_range[i])
  
  df_FF = rbind(df_FF,
                df_day)
  
  # show progress
  if (as.yearmon(date_range[i])!=prev_month) {
    print(as.yearmon(date_range[i]))
  }
  
  prev_month = as.yearmon(date_range[i])
}

save(df_FF, file = "df_FF.Rdata")
```

## Cleaning the data a little bit
```{r}
save(df_FF, file = "df_FF.Rdata")

df_FF_clean = df_FF %>%
  mutate(geo = NA,
         geo = ifelse(currency == "AUD", "Australia", geo),
         geo = ifelse(currency == "USD", "USA", geo),
         geo = ifelse(currency == "CHF", "Switzerland", geo),
         geo = ifelse(currency == "GBP", "UK", geo),
         geo = ifelse(currency == "EUR", "Euro Area", geo),
         geo = ifelse(currency == "CAD", "Canada", geo),
         geo = ifelse(currency == "NZD", "New Zealand", geo),
         geo = ifelse(currency == "CNY", "China", geo),
         geo = ifelse(grepl("French", event), "France", geo),
         geo = ifelse(grepl("Italian", event), "Italy", geo),
         geo = ifelse(grepl("Spanish", event), "Spain", geo),
         geo = ifelse(grepl("German", event), "German", geo)) %>%
  gather(-date, -geo, -currency, -event, key = var, value = value)

df_FF_clean %<>%
  group_by(date,geo,currency,event,var) %>%
  filter(n()>1)

df_FF_clean %<>%
  distinct() %>%
  mutate(value = str_remove(value, "%"),
         value = str_replace(value, "K", "000"),
         value = str_replace(value, "M", "000000"),
         value = str_replace(value, "B", "000000000"),
         value = as.numeric(value)) %>%
  spread(key = var, value = value) %>%
  relocate(date, geo, currency) %>%
  filter(!is.na(geo))

save(df_FF_clean, file = "df_FF_clean.Rdata")
```


# Building a full dataset excluding Forex Factory's details variables

## Installing the necessary python packages/libraries
```{r, eval = FALSE}
library(reticulate)
# py_install("undetected_chromedriver", pip = TRUE)
# py_install("selenium")
``` 

## Range of dates to build from
```{r}
date_range = seq(as.Date("2007-01-01"),as.Date("2022-02-06"), by = "days")
day_ref = day(date_range)
month_ref = str_to_lower(month(date_range, label = TRUE))
year_ref = year(date_range)
date_ref = paste0(month_ref, day_ref, ".", year_ref)
```

## Python code to save daily html pages with full details for each event
```{python}
import undetected_chromedriver as uc
from selenium import webdriver
from selenium.webdriver.common.by import By
import time
from selenium.webdriver.support.ui import WebDriverWait
from selenium.webdriver.support import expected_conditions as EC

# launching the webdriver
options = webdriver.ChromeOptions() 
options.add_argument("start-maximized")
driver = uc.Chrome(options=options)

# forex factor partial url
ff_url = 'https://www.forexfactory.com/calendar?day='

list_source = []

for day in r.date_ref:
  
  # url of the day
  ff_day_url = ff_url + day
  
  # go to the day url
  driver.get(ff_day_url)
  # time.sleep(2)

  # find events
  calandar_elem = driver.find_elements(By.CSS_SELECTOR, '[data-eventid].calendar_row')
  calandar_elem_open = driver.find_elements(By.CSS_SELECTOR, "[title='Open Detail']")
  
  # open details section for each event
  while len(calandar_elem_open)>0:
    for i in range(0,len(calandar_elem)):
      event = calandar_elem[i].get_attribute("data-eventid")
      
      if len(event)>0:
        css_sel = "[data-eventid='" + event + "'].calendar_row > .detail > [title='Open Detail']"
        
        details_elem = driver.find_element(By.CSS_SELECTOR, css_sel)
        driver.execute_script("arguments[0].click();", details_elem) 
        
    time.sleep(5)
    calandar_elem_open = driver.find_elements(By.CSS_SELECTOR, ".calendar_row > .detail > [title='Open Detail']")

  # store source html
  list_source.append(driver.page_source)
        
# driver.quit()
```

## saving html source pages
```{r}
full_html_source = py$list_source
save(full_html_source, file = "full_html_source.Rdata")
```

## Scrapping data from Forex Factory 
```{r}
load("full_html_source.Rdata")

df_FF_full = c()

prev_month = as.yearmon(date_range[1])

for (i in 1:length(full_html_source)) {
  
  elements <- read_html(full_html_source[i]) %>% 
    html_elements(".calendar__table") %>% 
    html_elements(".calendar__row--grey")
  
  if (length(elements) > 0) {
    
    currency <- elements %>%
      html_elements(".currency") %>%
      html_text2()
    
    event <- elements %>%
      html_elements(".event") %>%
      html_text2()
    
    forecast <- elements %>%
      html_elements(".forecast") %>%
      html_text2()
    
    observation <- elements %>%
      html_elements(".actual") %>%
      html_text2()
    
    previous_obs <- elements %>%
      html_elements(".previous") %>%
      html_text2()
    
    event_id = elements %>%
      html_attr("data-eventid")
    
    impact = elements %>%
      html_elements(".calendar__impact-icon--screen") %>%
      html_elements("[title]") %>%
      html_attr("title")
    
    # getting the details
    df_spec = c()
    
    for (id in event_id) {
      specs = read_html(full_html_source[i]) %>%
        html_element(paste0("[data-eventid='", id, "'].detail")) %>%
        html_element("tbody") %>%
        html_table() %>%
        mutate(event_id = id)
      
      df_spec = rbind(df_spec, specs)
    }
    
    df_spec %<>% 
      spread(key = X1, value = X2)
    ##
    
    df_day = tibble(currency,
                    event,
                    forecast,
                    observation,
                    previous_obs,
                    impact,
                    event_id) %>%
      mutate(date = date_range[i]) %>%
      full_join(df_spec, by = "event_id") %>%
      gather(-date, -event_id, key = var, value = value)
    
    df_FF_full = rbind(df_FF_full,
                       df_day)
    
  }
  
  # show progress
  if (as.yearmon(date_range[i])!=prev_month) {
    print(as.yearmon(date_range[i]))
  }
  
  prev_month = as.yearmon(date_range[i])
}

df_FF_full %<>%
  spread(key = var, value = value)

save(df_FF_full, file = "df_FF_full.Rdata")
```
      